---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cereal 🥣

<!-- badges: start -->
[![R-CMD-check](https://github.com/juliasilge/cereal/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/juliasilge/cereal/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/juliasilge/cereal/branch/main/graph/badge.svg)](https://app.codecov.io/gh/juliasilge/cereal?branch=main)
<!-- badges: end -->

The goal of cereal is to provide methods to [serialize](https://en.wikipedia.org/wiki/Serialization) objects from [vctrs](https://vctrs.r-lib.org/) to [JSON](https://en.wikipedia.org/wiki/JSON), as well as back from JSON to vctrs objects.

## Installation

You can install the development version of cereal from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("juliasilge/cereal")
```

## Example

There are many different data types that are useful in R analysis or models. In a dataframe or tibble, any single column has one type:

```{r example}
df <- tibble::tibble(
  a = c(1.2, 2.3, 3.4),
  b = 2L:4L,
  c = Sys.Date() + 0:2,
  d = as.POSIXct("2019-01-01", tz = "America/New_York") + 100:102,
  e = sample(letters, 3),
  f = factor(c("blue", "blue", "green"), levels = c("blue", "green", "red")),
  g = ordered(c("small", "large", "medium"), levels = c("small", "medium", "large"))
)

df
```

The vctrs package has a [concept of a **vector prototype**](https://vctrs.r-lib.org/articles/type-size.html) which captures the metadata associated with a vector without keeping any of the data itself.

```{r cars}
vctrs::vec_ptype(df)
```

The information stored in such a vector prototype includes, for example, the levels of a factor and the timezone for a datetime. This can be useful or important information when deploying code or models, such as when using [vetiver](https://vetiver.rstudio.com/). We could store this vector prototype as an R binary object saved as an `.rds` file, but with cereal, you can store this vector prototype in plain text as JSON:

```{r}
library(cereal)
json <- cereal_to_json(df, auto_unbox = TRUE, pretty = TRUE)
json
```

Storing prototype information as JSON (rather than a binary file) is more useful when it is needed as plain-text metadata, or anytime a human needs to read it. 

You can also convert from JSON back to the original prototype:

```{r}
cereal_from_json(json)
```

For an approach to this same task using Python, [see Pydantic's `model.json()`](https://docs.pydantic.dev/latest/usage/exporting_models/#modeljson).

## Contributing

- This project is released with a [Contributor Code of Conduct](https://www.contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

- If you think you have encountered a bug, please [submit an issue](https://github.com/juliasilge/cereal/issues).

- Either way, learn how to create and share a [reprex](https://reprex.tidyverse.org/articles/articles/learn-reprex.html) (a minimal, reproducible example), to clearly communicate about your code.


